---
title: "Analyzing the Dynamics of Home Advantage: A Statistical Investigation into Goals, Corners, and Shots on Target in Premier League Football"
author: 
  - Jerry Lu (Yen-Chia Lu)
thanks: "Code and data are available at: https://github.com/Jerryluuuu/-Premier-League.git"
date: 17 April, 2024
date-format: long
abstract: "This study explores the phenomenon of home advantage in the Premier League for the 2022–2023 campaign by analysing the correlation between goals, corners, and shots on target for the home club. Advanced statistical studies are performed using a season-long dataset to find trends and insights that clarify the significance of playing at home. Significant correlations that have practical consequences for clubs and coaching techniques as well as contributing to the scholarly discourse on sports analytics are highlighted by the findings. In light of the most recent season, this research provides fresh perspectives on the tactical and psychological factors affecting football game results."
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
####workspcae setup####
library(tidyverse)
library(here)
library(rstanarm)
library(modelsummary)
library(knitr)
library(dplyr)
library(tibble)
```

```{r}
#| include: false
#| warning: false
#| message: false
#### import data ####
analysis_PL_data <- read_csv(here::here("data/analysis_data/analysis_PL_data.csv"))
```


# Introduction

Since its founding in 1992, the Premier League has grown to become one of the world's most competitive and watched football competitions. Each of the twenty teams that compete each year plays 38 games, half of which are played at their home stadium and the other half away. The league is renowned for its competitive matches, fervent fandoms, and the exceptional skill level of its clubs and players. This competitive atmosphere creates an ideal setting for investigating several aspects of a team's performance, especially the home field advantage.

In sports, the idea of home team advantage is widely accepted, and the Premier League is no different. Playing at home is frequently linked to a higher chance of success, which can be related to things like comfortable surroundings, the backing of local supporters, and the lack of travel fatigue. Metrics that are essential for assessing a team's attacking prowess at home games include Full Time Home Team Goals, Home Team Shots on Target (HST), and Home Team Corners (HC). Understanding how home advantage might affect a team's attacking dynamics and overall performance in a match depends on these factors.

Not to be overlooked in this situation is the referee's involvement. Although it is the referees' responsibility to maintain fair play, their judgements have a big influence on how the game goes and how it ends. There is interest in the idea of "home bias," which holds that choices may unintentionally benefit the home team, for example, by giving them extra time off from injuries or more lenient rulings for fouls. Understanding the psychological effects of home advantage and the impact of outside variables on game outcomes can be gained by examining the relationship between referee decisions and home team performance indicators like goals, shots on target, and corners.

Furthermore, a key factor in maximising home advantage is the home team's confidence. When combined with the encouragement of home fans, the psychological boost of playing in a familiar environment can improve team performance. Higher percentages of shots on target and in corners are signs of an aggressive offensive strategy that frequently reflects this confidence. We can measure the effect of home confidence on a team's performance by looking at these variables, which provides a greater understanding of the psychological and strategic components that contribute to the Premier League's home advantage occurrences. 

# Data {#sec-data}

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.


```{r}
#| label: tbl-PLhome
#| tbl-cap: "Premier League Home Team data in 2023 season"
#| echo: false

Home_team_averages <- analysis_PL_data |>
  group_by(`Home Team`) |>
  summarise(Avg_FTHG = mean(`Full Time Home Team Goals`), Avg_HC = mean(`Home Team Corners`), Avg_HST = mean(`Home Team Shots on Target`)) |>
  arrange(desc(Avg_FTHG)) |>
  kable(
    col.names = c("Team", "Avg FTHG", "Avg HC", "Avg HST"),
    digits = 1
  )
print(Home_team_averages)
```

```{r}
#| label: tbl-PLaway
#| tbl-cap: "Premier League Away Team data in 2023 season"
#| echo: false

Away_team_averages <- analysis_PL_data |>
  group_by(`Away Team`) |>
  summarise(Avg_FTAG = mean(`Full Time Away Team Goals`), Avg_AC = mean(`Away Team Corners`), Avg_AST = mean(`Away Team Shots on Target`)) |>
  arrange(desc(Avg_FTAG)) |>
  kable(
    col.names = c("Team", "Avg FTAG", "Avg AC", "Avg AST"),
    digits = 1
  )
print(Away_team_averages)
```

```{r}
#| label: fig-PLHChome
#| tbl-cap: "Home Coners and Win rate relationship"
#| echo: false
#| message: false

team_stats <- analysis_PL_data |>
  group_by(`Home Team`) |>
  summarise(
    Avg_HC = mean(`Home Team Corners`),
    Home_Wins = sum(`Full Time Result` == 'H'),
    Games_Played = n(),
    Home_Win_Rate = Home_Wins / Games_Played * 100
  )
ggplot(team_stats, aes(x = Avg_HC, y = Home_Win_Rate)) +
  geom_point(aes(color = `Home Team`), size = 3) +  # Color coded by team
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add a linear regression line without confidence interval
  labs(title = "Relationship Between Home Corners and Home Win Rate",
       x = "Average Home Corners",
       y = "Home Win Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-PLHCaway
#| tbl-cap: "Away Coners and Win rate relationship"
#| echo: false
#| message: false

team_stats_away <- analysis_PL_data %>%
  group_by(`Away Team`) %>%
  summarise(
    Avg_AC = mean(`Away Team Corners`),
    Away_Wins = sum(`Full Time Result` == 'A'),
    Games_Played = n(),
    Away_Win_Rate = Away_Wins / Games_Played * 100
  )

ggplot(team_stats_away, aes(x = Avg_AC, y = Away_Win_Rate)) +
  geom_point(aes(color = `Away Team`), size = 3) +  # Color coded by team
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without confidence interval
  labs(title = "Relationship Between Away Corners and Away Win Rate",
       x = "Average Away Corners",
       y = "Away Win Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-PLHSThome
#| tbl-cap: "Home Team Shot on Target and Win rate relationship"
#| echo: false
#| message: false
team_stats_home_hst <- analysis_PL_data %>%
  group_by(`Home Team`) %>%
  summarise(
    Avg_HST = mean(`Home Team Shots on Target`, na.rm = TRUE),
    Home_Wins = sum(`Full Time Result` == 'H'),
    Games_Played = n(),
    Home_Win_Rate = Home_Wins / Games_Played * 100
  )
ggplot(team_stats_home_hst, aes(x = Avg_HST, y = Home_Win_Rate)) +
  geom_point(aes(color = `Home Team`), size = 3) +  # Color coded by team
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add a linear regression line without confidence interval
  labs(title = "Relationship Between Home Shots on Target and Home Win Rate",
       x = "Average Home Shots on Target",
       y = "Home Win Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-PLHSTaway
#| tbl-cap: "Away Team Shot on Target and Win rate relationship"
#| echo: false
#| message: false
team_stats_away_ast <- analysis_PL_data %>%
  group_by(`Away Team`) %>%
  summarise(
    Avg_AST = mean(`Away Team Shots on Target`, na.rm = TRUE),  
    Away_Wins = sum(`Full Time Result` == 'A'),
    Games_Played = n(),
    Away_Win_Rate = Away_Wins / Games_Played * 100
  )
ggplot(team_stats_away_ast, aes(x = Avg_AST, y = Away_Win_Rate)) +
  geom_point(aes(color = `Away Team`), size = 3) +  # Color coded by team
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without confidence interval
  labs(title = "Relationship Between Away Shots on Target and Away Win Rate",
       x = "Average Away Shots on Target",
       y = "Away Win Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-PLhomevsaway
#| tbl-cap: "Home ad Away win rates for each team in 2023 season"
#| echo: false
#| message: false
home_win_rates <- analysis_PL_data |>
  group_by(`Home Team`) |>
  summarise(
    Home_Wins = sum(`Full Time Result` == 'H'),
    Home_Games = n(),
    Home_Win_Rate = Home_Wins / Home_Games * 100
  ) |>
  rename(Team = `Home Team`)
away_win_rates <- analysis_PL_data |>
  group_by(`Away Team`) |>
  summarise(
    Away_Wins = sum(`Full Time Result` == 'A'),
    Away_Games = n(),
    Away_Win_Rate = Away_Wins / Away_Games * 100
  ) |>
  rename(Team = `Away Team`)
combined_win_rates <- full_join(home_win_rates, away_win_rates, by = "Team")
ggplot(combined_win_rates, aes(x = Team)) +
  geom_bar(aes(y = Home_Win_Rate, fill = "Home"), stat = "identity", position = position_dodge(width = 0.9)) +
  geom_bar(aes(y = Away_Win_Rate, fill = "Away"), stat = "identity", position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("Home" = "green", "Away" = "purple")) +
  labs(title = "Home and Away Win Rates for Each Team", x = "Team", y = "Win Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

```{r}
#| label: fig-PLrefuree
#| tbl-cap: "The relationship between refuree and win rate"
#| echo: false
#| message: false
overall_home_win_rate <- mean(analysis_PL_data$`Full Time Result` == 'H')
overall_away_win_rate <- mean(analysis_PL_data$`Full Time Result` == 'A')

referee_influence <- analysis_PL_data |>
  group_by(Referee) |>
  summarise(
    Home_Wins = sum(`Full Time Result` == 'H'),
    Away_Wins = sum(`Full Time Result` == 'A'),
    Games = n(),
    Home_Win_Rate = Home_Wins / Games * 100,
    Away_Win_Rate = Away_Wins / Games * 100,
    Deviation_Home = Home_Win_Rate - overall_home_win_rate * 100,
    Deviation_Away = Away_Win_Rate - overall_away_win_rate * 100
  )
ggplot(referee_influence, aes(x = Referee)) +
  geom_bar(aes(y = Deviation_Home, fill = "Home"), stat = "identity", position = position_dodge()) +
  geom_bar(aes(y = Deviation_Away, fill = "Away"), stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("Home" = "blue", "Away" = "red")) +
  labs(title = "Referee Influence on Home and Away Win Rates", x = "Referee", y = "Deviation from Average Win Rate (%)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

# Model

## Model set-up

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

PL_data_first_model_rstanarm <-
  readRDS(file = here::here("models/PL_model.rds"))
```

```{r}
#| label: fig-models
#| fig-cap: "The relationship of Full Time Home Team Goals on the Home Team shot on Target and Home Corners"  
#| echo: false
#| warning: false
#| message: false

draws <- 1000  # Number of simulations per shot on target

# Create a tibble with proper length matching
analysis_PL_data <-
  tibble(
    sigma = rep(rexp(n = draws, rate = 1), each = 11),  # Repeat each sigma value 11 times
    beta_0 = rep(rnorm(n = draws, mean = 0, sd = 2.5), each = 11),
    beta_1 = rep(rnorm(n = draws, mean = 15, sd = 2.5), each = 11),
    `Home Team Shots on Target` = rep(0:10, times = draws),  # Ensure this matches the 11000 length
    mu = beta_0 + beta_1 * `Home Team Shots on Target`
  ) %>%
  rowwise() %>%
  mutate(
    `Full Time Home Team Goals` = rnorm(n = 1, mean = mu, sd = sigma)
  ) %>%
  ungroup()  # Ensure data is ungrouped for plotting

# Plotting the relationship
ggplot(analysis_PL_data, aes(x = `Home Team Shots on Target`, y = `Full Time Home Team Goals`)) +
  geom_point(alpha = 0.2, color = "blue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
  labs(title = "Relationship Between Shots on Target and Goals",
       x = "Home Team Shots on Target",
       y = "Full Time Home Team Goals")
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of Premier League Full Time Home Team Goals on Home Team Shots on Target and Home Team Corners"
#| warning: false

modelsummary(PL_data_first_model_rstanarm)
```


# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check


\newpage


# References


